Random number generation and applications on GPU’s in R?

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
back to previous version of fisher
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2873 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Comparing running time: Month data example}
<<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
data2<-read.csv("/home/ruoyong/paper1_2021_5/month.csv") 
data2<-data2[,-1]
month<-as.matrix(data2)
rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
Nglobal = c(128,128)
streams <- CreateStreamsGpu(Nstreams =128*128, keepInitial=1)
@
<<monthdata,eval=TRUE,echo=FALSE, cache=FALSE>>=
library(knitr)
library(kableExtra)
knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@

@@

month_GPU<-vclMatrix(month,type="integer")
result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
result_monthgpu$simNum
result_monthgpu$counts
result_monthgpu$p.value
@
We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
<<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
#using GPU
system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
@
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU. 
<<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
result_monthgpu$cpu <- as.vector(result_monthgpu$sim)
@

@@

<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@
We perform 10 million simulations for the ``week'' table.
<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
#using CPU
fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
@
The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
<<TimecompareweekGpu2, eval=TRUE, cache=TRUE>>=
observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1

week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = Nglobal)
result_weekgpu$simNum
result_weekgpu$counts
result_weekgpu$p.value
@
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
<<TimecompareweekGpu, eval=TRUE, cache=TRUE>>=
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
@
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
result_weekgpu$cpu = as.vector(result_weekgpu$sim)
@

>>>>

\subsection{Comparing running time: Month data example}
<<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
data2<-read.csv("/home/ruoyong/paper1_2021_5/month.csv") 
data2<-data2[,-1]
month<-as.matrix(data2)
rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
Nglobal = c(512,128)
streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
@
<<monthdata,eval=TRUE,echo=FALSE, cache=FALSE>>=
library(knitr)
library(kableExtra)
knitr::kable(month,format="latex",  caption = "Month\\label{tab:month}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@

@@

month_GPU<-vclMatrix(month,type="integer")
result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(512,128))
result_monthgpu$simNum
result_monthgpu$counts
result_monthgpu$p.value
@
We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
<<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
#using GPU
system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(512,128)))
@
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU. 
<<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
result_monthgpu$cpu <- as.vector(result_monthgpu$sim)
@

@@

<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",caption = "Week\\label{tab:week}") %>%
  kable_styling(full_width = F, position = "center",latex_options = "HOLD_position")
@
We perform 10 million simulations for the ``week'' table.
<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
#using CPU
fisher.test(week,simulate.p.value = TRUE,B=1e7)$p.value
system.time(fisher.test(week,simulate.p.value = TRUE,B=1e7))
@
The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.
<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1

week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = Nglobal)
result_weekgpu$simNum
result_weekgpu$counts
result_weekgpu$p.value
@
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
<<TimecompareweekGpu, eval=TRUE, cache=FALSE>>=
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
@
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU. 
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
result_weekgpu$cpu = as.vector(result_weekgpu$sim)
@

### 2021-09-28 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
back to previous version of fisher
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2873 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

month_GPU<-vclMatrix(month,type="integer")
result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(128,128))
result_monthgpu$simNum
## [1] 1000064
result_monthgpu$counts
## [1] 403788
result_monthgpu$p.value
## [1] 0.4037628
\end{verbatim}
\end{kframe}
\end{knitrout}
We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(128,128)))
##    user  system elapsed 
##   0.793   0.000   0.792
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.

@@

week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = Nglobal)
result_weekgpu$simNum
## NULL
result_weekgpu$counts
## [1] 1224
result_weekgpu$p.value
## [1] 0.0001225
\end{verbatim}
\end{kframe}
\end{knitrout}
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
##    user  system elapsed 
##   1.899   0.000   1.898
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.

>>>>

month_GPU<-vclMatrix(month,type="integer")
result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(512,128))
result_monthgpu$simNum
## [1] 1048576
result_monthgpu$counts
## [1] 422440
result_monthgpu$p.value
## [1] 0.4028707
\end{verbatim}
\end{kframe}
\end{knitrout}
We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(512,128)))
##    user  system elapsed 
##   0.719   0.011   0.730
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.

@@

week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = Nglobal)
result_weekgpu$simNum
## [1] 10027008
result_weekgpu$counts
## [1] 1226
result_weekgpu$p.value
## [1] 0.0001223695
\end{verbatim}
\end{kframe}
\end{knitrout}
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = Nglobal))
##    user  system elapsed 
##   4.649   0.004   4.710
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.
